---
title: "Adding annotation information to beta values"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

## Setup

Loading packages

```{r, message=FALSE, warning=FALSE}
library(sesame)
library(tidyverse)
library(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
```

```{r}
idat_folder <- "~/Data/methylation/idat"
```

Example of extraction of beta values for 1 sample using **sesame**.

```{r}
sesame_data <- openSesame(file.path(idat_folder, "200514030126_R06C01"))
```

we then turn it into a tibble with two columns, `beta` and `site`

```{r}
sesame_df <- as.data.frame(sesame_data) %>%
  rownames_to_column(var = "site") %>%
  dplyr::rename(beta = sesame_data)

head(sesame_df)
```

Here we create a tibble of chromosome location information

```{r}
chr_ref <- tibble(
  chr = Locations$chr,
  site = Locations@rownames
)

head(chr_ref)
```

lastly we can join those two tibbles together

```{r}
sesame_df_chr <- sesame_df %>%
  left_join(chr_ref, by = "site")

head(sesame_df_chr)
```

This is a general workflow, `IlluminaHumanMethylationEPICanno.ilm10b4.hg19` includes many other annotation information that can be used.